---
title: "Charlottesville/Albemarle Schools"
author: "Compiled by Michele Claibourn, Nina Schoonover, & Elizabeth Mitchell, UVA Equity Center"
date: "3/1/2023"
output: html_document
subtitle: Demographic Composition
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(reactable)

load("newstudents.Rdata")
students2023 <- students2023 %>% 
  separate(division, c("division", "locality")) %>% 
  select(-c(year, locality)) %>% 
  relocate(c(count_adv, count_disadv), .before = perc_aian)
students2022 <- students2022 %>% 
  separate(division, c("division", "locality")) %>% 
  select(-c(year, locality)) %>% 
  relocate(c(count_adv, count_disadv), .before = perc_aian)
students2021 <- students2021 %>% 
  separate(division, c("division", "locality")) %>% 
  select(-c(year, locality)) %>% 
  relocate(c(count_adv, count_disadv), .before = perc_aian)
students2020 <- students2020 %>% 
  separate(division, c("division", "locality")) %>% 
  select(-c(year, locality)) %>% 
  relocate(c(count_adv, count_disadv), .before = perc_aian)

```

## Demographic Tables {.tabset}
The tables below provide the demographic composition of Charlottesville and Albemarle Schools for 2022-2023, 2021-2022, 2020-2021, and 2019-2020.

These numbers reflect Fall Membership numbers provided by the Virginia Department of Education's [Fall Membership Build-a-Table site](https://p1pe.doe.virginia.gov/apex_captcha/home.do?apexTypeId=304).

According to VDOE documentation

* Race or Ethnicity captures "the racial category that most closely reflects the student’s recognition of his or her community or with which the student identifies"
* Economic status "identifies students as economically disadvantaged if they meet any one of the following: 1) is eligible for Free/Reduced Meals, or 2) receives TANF, or 3) is eligible for Medicaid, or 4) identified as either Migrant or experiencing Homelessness."

Scroll further right to see the percent of a school's student body by race and the number and percent of students identified as economically disadvantaged. The tables can be sorted by any column by clicking on the up/down arrow next to the column name.

### 2022-2023 Academic Year
```{r}
reactable(students2023,
          columns = list(
             division = colDef(name = "Division",
                                style = list(background = "#F4E3D3FF")),
            school = colDef(name = "School",
                                style = list(background = "#F4E3D3FF")),
            students = colDef(name = "Students",
                                style = list(background = "#F4E3D3FF")),
            count_aian = colDef(name = "Am. Indian AK Native",
                                style = list(background = "#C5CCD2")),
            count_asian = colDef(name = "Asian",
                                style = list(background = "#C5CCD2")),
            count_black = colDef(name = "Black",
                                style = list(background = "#C5CCD2")),
            count_latinx = colDef(name = "Hispanic",
                                style = list(background = "#C5CCD2")),
            count_nhpi = colDef(name = "Native HI Pacific Islander",
                                style = list(background = "#C5CCD2")),
            count_white = colDef(name = "White",
                                style = list(background = "#C5CCD2")),
            count_adv = colDef(name = "Advantaged",
                                style = list(background = "#B4D6DB")),
            count_disadv = colDef(name = "Disadvantaged",
                                style = list(background = "#B4D6DB")),
            perc_aian = colDef(name = "Am. Indian AK Native",
                               style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_asian = colDef(name = "Asian",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_black = colDef(name = "Black",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_latinx = colDef(name = "Hispanic",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_nhpi = colDef(name = "Native HI Pacific Islander",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_white = colDef(name = "White",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_adv = colDef(name = "Advantaged",
                                style = list(background = "#94C5CCFF"),
                               format = colFormat(suffix = "%")),
            perc_disadv = colDef(name = "Disadvantaged",
                                style = list(background = "#94C5CCFF"),
                               format = colFormat(suffix = "%"))),
          columnGroups = list(
            colGroup(name = "Counts by Racial/Ethnic Identity", 
                     columns = c("count_aian", "count_asian", "count_black", "count_latinx",
                                 "count_nhpi", "count_white")),
            colGroup(name = "Counts by Economic Status", 
                     columns = c("count_adv", "count_disadv")),
            colGroup(name = "Percents by Racial/Ethnic Identity", 
                     columns = c("perc_aian", "perc_asian", "perc_black", "perc_latinx",
                                 "perc_nhpi", "perc_white")),
            colGroup(name = "Percents by Economic Status", 
                     columns = c("perc_adv", "perc_disadv"))
          ),
          showSortable = TRUE, defaultPageSize = 12)
          
```

### 2021-2022 Academic Year
```{r}
reactable(students2022,
          columns = list(
            division = colDef(name = "Division",
                                style = list(background = "#F4E3D3FF")),
            school = colDef(name = "School",
                                style = list(background = "#F4E3D3FF")),
            students = colDef(name = "Students",
                                style = list(background = "#F4E3D3FF")),
            count_aian = colDef(name = "Am. Indian AK Native",
                                style = list(background = "#C5CCD2")),
            count_asian = colDef(name = "Asian",
                                style = list(background = "#C5CCD2")),
            count_black = colDef(name = "Black",
                                style = list(background = "#C5CCD2")),
            count_latinx = colDef(name = "Hispanic",
                                style = list(background = "#C5CCD2")),
            count_nhpi = colDef(name = "Native HI Pacific Islander",
                                style = list(background = "#C5CCD2")),
            count_white = colDef(name = "White",
                                style = list(background = "#C5CCD2")),
            count_adv = colDef(name = "Advantaged",
                                style = list(background = "#B4D6DB")),
            count_disadv = colDef(name = "Disadvantaged",
                                style = list(background = "#B4D6DB")),
            perc_aian = colDef(name = "Am. Indian AK Native",
                               style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_asian = colDef(name = "Asian",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_black = colDef(name = "Black",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_latinx = colDef(name = "Hispanic",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_nhpi = colDef(name = "Native HI Pacific Islander",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_white = colDef(name = "White",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_adv = colDef(name = "Advantaged",
                                style = list(background = "#94C5CCFF"),
                               format = colFormat(suffix = "%")),
            perc_disadv = colDef(name = "Disadvantaged",
                                style = list(background = "#94C5CCFF"),
                               format = colFormat(suffix = "%"))),
          columnGroups = list(
            colGroup(name = "Counts by Racial/Ethnic Identity", 
                     columns = c("count_aian", "count_asian", "count_black", "count_latinx",
                                 "count_nhpi", "count_white")),
            colGroup(name = "Counts by Economic Status", 
                     columns = c("count_adv", "count_disadv")),
            colGroup(name = "Percents by Racial/Ethnic Identity", 
                     columns = c("perc_aian", "perc_asian", "perc_black", "perc_latinx",
                                 "perc_nhpi", "perc_white")),
            colGroup(name = "Percents by Economic Status", 
                     columns = c("perc_adv", "perc_disadv"))
          ),
          showSortable = TRUE, defaultPageSize = 12)
```

### 2020-2021 Academic Year
```{r}
reactable(students2021,
          columns = list(
            division = colDef(name = "Division",
                                style = list(background = "#F4E3D3FF")),
            school = colDef(name = "School",
                                style = list(background = "#F4E3D3FF")),
            students = colDef(name = "Students",
                                style = list(background = "#F4E3D3FF")),
            count_aian = colDef(name = "Am. Indian AK Native",
                                style = list(background = "#C5CCD2")),
            count_asian = colDef(name = "Asian",
                                style = list(background = "#C5CCD2")),
            count_black = colDef(name = "Black",
                                style = list(background = "#C5CCD2")),
            count_latinx = colDef(name = "Hispanic",
                                style = list(background = "#C5CCD2")),
            count_nhpi = colDef(name = "Native HI Pacific Islander",
                                style = list(background = "#C5CCD2")),
            count_white = colDef(name = "White",
                                style = list(background = "#C5CCD2")),
            count_adv = colDef(name = "Advantaged",
                                style = list(background = "#B4D6DB")),
            count_disadv = colDef(name = "Disadvantaged",
                                style = list(background = "#B4D6DB")),
            perc_aian = colDef(name = "Am. Indian AK Native",
                               style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_asian = colDef(name = "Asian",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_black = colDef(name = "Black",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_latinx = colDef(name = "Hispanic",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_nhpi = colDef(name = "Native HI Pacific Islander",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_white = colDef(name = "White",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_adv = colDef(name = "Advantaged",
                                style = list(background = "#94C5CCFF"),
                               format = colFormat(suffix = "%")),
            perc_disadv = colDef(name = "Disadvantaged",
                                style = list(background = "#94C5CCFF"),
                               format = colFormat(suffix = "%"))),
          columnGroups = list(
            colGroup(name = "Counts by Racial/Ethnic Identity", 
                     columns = c("count_aian", "count_asian", "count_black", "count_latinx",
                                 "count_nhpi", "count_white")),
            colGroup(name = "Counts by Economic Status", 
                     columns = c("count_adv", "count_disadv")),
            colGroup(name = "Percents by Racial/Ethnic Identity", 
                     columns = c("perc_aian", "perc_asian", "perc_black", "perc_latinx",
                                 "perc_nhpi", "perc_white")),
            colGroup(name = "Percents by Economic Status", 
                     columns = c("perc_adv", "perc_disadv"))
          ),
          showSortable = TRUE, defaultPageSize = 13)
```

### 2019-2020 Academic Year
```{r}
reactable(students2020,
          columns = list(
            division = colDef(name = "Division",
                                style = list(background = "#F4E3D3FF")),
            school = colDef(name = "School",
                                style = list(background = "#F4E3D3FF")),
            students = colDef(name = "Students",
                                style = list(background = "#F4E3D3FF")),
            count_aian = colDef(name = "Am. Indian AK Native",
                                style = list(background = "#C5CCD2")),
            count_asian = colDef(name = "Asian",
                                style = list(background = "#C5CCD2")),
            count_black = colDef(name = "Black",
                                style = list(background = "#C5CCD2")),
            count_latinx = colDef(name = "Hispanic",
                                style = list(background = "#C5CCD2")),
            count_nhpi = colDef(name = "Native HI Pacific Islander",
                                style = list(background = "#C5CCD2")),
            count_white = colDef(name = "White",
                                style = list(background = "#C5CCD2")),
            count_adv = colDef(name = "Advantaged",
                                style = list(background = "#B4D6DB")),
            count_disadv = colDef(name = "Disadvantaged",
                                style = list(background = "#B4D6DB")),
            perc_aian = colDef(name = "Am. Indian AK Native",
                               style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_asian = colDef(name = "Asian",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_black = colDef(name = "Black",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_latinx = colDef(name = "Hispanic",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_nhpi = colDef(name = "Native HI Pacific Islander",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_white = colDef(name = "White",
                                style = list(background = "#ADB7C0FF"),
                               format = colFormat(suffix = "%")),
            perc_adv = colDef(name = "Advantaged",
                                style = list(background = "#94C5CCFF"),
                               format = colFormat(suffix = "%")),
            perc_disadv = colDef(name = "Disadvantaged",
                                style = list(background = "#94C5CCFF"),
                               format = colFormat(suffix = "%"))),
          columnGroups = list(
            colGroup(name = "Counts by Racial/Ethnic Identity", 
                     columns = c("count_aian", "count_asian", "count_black", "count_latinx",
                                 "count_nhpi", "count_white")),
            colGroup(name = "Counts by Economic Status", 
                     columns = c("count_adv", "count_disadv")),
            colGroup(name = "Percents by Racial/Ethnic Identity",
                     columns = c("perc_aian", "perc_asian", "perc_black", "perc_latinx",
                                 "perc_nhpi", "perc_white")),
            colGroup(name = "Percents by Economic Status", 
                     columns = c("perc_adv", "perc_disadv"))
          ),
          showSortable = TRUE, defaultPageSize = 13)
```


