---
title: "Advanced Diplomas"
date: "2024-03-28"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(janitor)
library(scales)
library(plotly)
```

```{r dataprep}
advanced3 <- read_csv("advanced_diplomas_3years.csv")

adv23 <- advanced3 %>% 
  select(year, division, starts_with("percent")) %>% 
  filter(year == 2022) %>% 
  mutate(division = str_remove(division, " Public Schools"),
         order = dense_rank(`percent_advanced_All Students`),
         highlight = case_when(
           division == "Albemarle County" ~ "alb", 
           division == "Charlottesville City" ~ "cvl", 
           TRUE ~ "other")) %>% 
  pivot_longer(starts_with("percent"), names_prefix = "percent_advanced_",
               names_to = "group", values_to = "advanced") %>% 
  mutate(advanced = round(advanced, 1))
```


## Advanced Diplomas, 2022-2023 Class

These data represent the percent of students graduating from each Virginia school division with an advanced diploma in 2023, as provided in the [School Quality Profiles](https://schoolquality.virginia.gov/).[^1]  

The School Quality Profiles page offers the following description of diplomas and school completion outcomes:

> Most Virginia students earn either an Advanced Studies Diploma or a Standard Diploma.

> To graduate with an Advanced Studies Diploma, a student must earn at least 26 standard units of credit by passing required courses and electives and at least nine verified units of credit by passing Standards of Learning end-of-course assessments in English, mathematics, science and history. Students who entered the ninth grade in 2013-2014 and afterwards must also successfully complete one virtual course.

> To graduate with a Standard Diploma, a student must earn at least 22 standard units of credit by passing required courses and electives, and earn at least six verified credits by passing end-of-course SOL tests or other assessments approved by the Board of Education. Students who entered the ninth grade in 2013-2014 and afterwards must earn a board-approved career and technical education credential to graduate and successfully complete one virtual course.

### Among All Students (%)

```{r class23, fig.height=10}
p <- adv23 %>% 
  filter(group == "All Students") %>% 
  ggplot(aes(y = fct_reorder(division, order), x = advanced, 
             color = highlight, 
             text = paste("Division:", division))) +
  geom_point(aes(size = highlight)) +
  geom_segment(aes(y = division, yend = division,
                   x = 0, xend = advanced)) +
  labs(x = "Percent Advanced Diplomas", y = "") +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  scale_color_manual(values = c("red", "orange", "grey"),
                     guide = "none") +
  scale_size_manual(values = c(3,3,2),
                    guide = "none") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 7),
        legend.position = "none") 

ggplotly(p, tooltip = c("text", "x"))

```

### By Race or Ethnicity

```{r race_white, fig.height = 10}
# White students (%)
p1 <- advanced3 %>% 
  select(year, division, percent_advanced_White) %>% 
  filter(year == 2022, !is.na(percent_advanced_White)) %>% 
  mutate(division = str_remove(division, " Public Schools"),
         order = dense_rank(percent_advanced_White),
         highlight = case_when(
           division == "Albemarle County" ~ "alb", 
           division == "Charlottesville City" ~ "cvl", 
           TRUE ~ "other")) %>% 
  pivot_longer(starts_with("percent"), names_prefix = "percent_advanced_",
               names_to = "group", values_to = "advanced") %>% 
  mutate(advanced = round(advanced, 1)) %>% 
  ggplot(aes(y = fct_reorder(division, order), x = advanced, 
             color = highlight,
             text = paste("Division:", division))) +
  geom_point(aes(size = highlight)) +
  geom_segment(aes(y = division, yend = division,
                   x = 0, xend = advanced)) +
  labs(x = "Percent Advanced Diplomas", y = "", title = "Among White Students") +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  scale_color_manual(values = c("red", "orange", "grey"),
                     guide = "none") +
  scale_size_manual(values = c(3,3,2),
                    guide = "none") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 7),
        legend.position = "none")

ggplotly(p1, tooltip = c("text", "x"))
```

```{r race_black, fig.height = 8}
# Black students (%)
p2 <- advanced3 %>% 
  select(year, division, percent_advanced_Black) %>% 
  filter(year == 2022, !is.na(percent_advanced_Black)) %>% 
  mutate(division = str_remove(division, " Public Schools"),
         order = dense_rank(percent_advanced_Black),
         highlight = case_when(
           division == "Albemarle County" ~ "alb", 
           division == "Charlottesville City" ~ "cvl", 
           TRUE ~ "other")) %>% 
  pivot_longer(starts_with("percent"), names_prefix = "percent_advanced_",
               names_to = "group", values_to = "advanced") %>% mutate(advanced = round(advanced, 1)) %>% 
  ggplot(aes(y = fct_reorder(division, order), x = advanced, 
             color = highlight,
             text = paste("Division:", division))) +
  geom_point(aes(size = highlight)) +
  geom_segment(aes(y = division, yend = division,
                   x = 0, xend = advanced)) +
  labs(x = "Percent Advanced Diplomas", y = "", title = "Among Black Students") +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  scale_color_manual(values = c("red", "orange", "grey"),
                     guide = "none") +
  scale_size_manual(values = c(3,3,2),
                    guide = "none") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 7),
        legend.position = "none")

ggplotly(p2, tooltip = c("text", "x"))
```

```{r race_hisp, fig.height = 7}
# Hispanic students (%)
p3 <- advanced3 %>% 
  select(year, division, percent_advanced_Hispanic) %>% 
  filter(year == 2022, !is.na(percent_advanced_Hispanic)) %>% 
  mutate(division = str_remove(division, " Public Schools"),
         order = dense_rank(percent_advanced_Hispanic),
         highlight = case_when(
           division == "Albemarle County" ~ "alb", 
           division == "Charlottesville City" ~ "cvl", 
           TRUE ~ "other")) %>% 
  pivot_longer(starts_with("percent"), names_prefix = "percent_advanced_",
               names_to = "group", values_to = "advanced") %>% mutate(advanced = round(advanced, 1)) %>% 
  ggplot(aes(y = fct_reorder(division, order), x = advanced, 
             color = highlight,
             text = paste("Division:", division))) +
  geom_point(aes(size = highlight)) +
  geom_segment(aes(y = division, yend = division,
                   x = 0, xend = advanced)) +
  labs(x = "Percent Advanced Diplomas", y = "", title = "Among Hispanic Students") +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  scale_color_manual(values = c("red", "orange", "grey"),
                     guide = "none") +
  scale_size_manual(values = c(3,3,2),
                    guide = "none") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 7),
        legend.position = "none")

ggplotly(p3, tooltip = c("text", "x"))
```

```{r race_asian, fig.height = 4}
# Asian students (%)
p4 <- advanced3 %>% 
  select(year, division, percent_advanced_Asian) %>% 
  filter(year == 2022, !is.na(percent_advanced_Asian)) %>% 
  mutate(division = str_remove(division, " Public Schools"),
         order = dense_rank(percent_advanced_Asian),
         highlight = case_when(
           division == "Albemarle County" ~ "alb", 
           division == "Charlottesville City" ~ "cvl", 
           TRUE ~ "other")) %>% 
  pivot_longer(starts_with("percent"), names_prefix = "percent_advanced_",
               names_to = "group", values_to = "advanced") %>% mutate(advanced = round(advanced, 1)) %>% 
  ggplot(aes(y = fct_reorder(division, order), x = advanced, 
             color = highlight,
             text = paste("Division:", division))) +
  geom_point(aes(size = highlight)) +
  geom_segment(aes(y = division, yend = division,
                   x = 0, xend = advanced)) +
  labs(x = "Percent Advanced Diplomas", y = "", title = "Among Asian Students") +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  scale_color_manual(values = c("red", "orange", "grey"),
                     guide = "none") +
  scale_size_manual(values = c(3,3,2),
                    guide = "none") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 7),
        legend.position = "none")

ggplotly(p4, tooltip = c("text", "x"))
```

## Advanced Diplomas, 2020-21 through 2022-23 Class

Below are the same data for the three years for which data are available to download from the [School Quality Profiles pages](https://schoolquality.virginia.gov/download-data), shown for only a subset of school divisions.

### Among All Students

In selected jurisdictions

```{r class2123}
adv2123 <- advanced3 %>% 
  select(year, division, starts_with("percent")) %>% 
  filter(str_detect(division, 
                    "Loudoun|Fairfax|Chesterfield|
                    Fluvanna|Prince William|Louisa|
                    Orange|Greene|Albemarle|
                    Virginia Beach|Charlottesville|Henrico|
                    Buckingham|Richmond City|Waynesboro")) %>% 
  mutate(division = str_remove(division, " Public Schools"),
         order = dense_rank(`percent_advanced_All Students`),
         highlight = ifelse(division == "Albemarle County", "yes", "no")) %>% 
  pivot_longer(starts_with("percent"), names_prefix = "percent_advanced_",
               names_to = "group", values_to = "advanced") %>% mutate(advanced = round(advanced, 1))

adv2123_endpoints <- adv2123 %>% 
  filter(group == "All Students", year == 2022)

p <- adv2123 %>% 
  filter(group == "All Students") %>% 
  ggplot(aes(y = advanced, x = year, color = as.factor(division), text = paste("Division:", division))) +
  geom_line() +
  geom_point() +
  geom_text(data = adv2123_endpoints, size = 4,
            aes(label = division, x = year + 0.3)) + 
  labs(y = "Percent Advanced Diplomas", x = "") +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  scale_x_continuous(breaks = c(2020, 2021, 2022),
                     expand = expansion(add = c(0.05,0.33))) +
  scale_color_manual(values = c("red", "orange", rep("grey",9)),
                     guide = "none") +
  theme_minimal() +
  theme(legend.position = "none")

ggplotly(p, tooltip = c("text", "y", "x"))
```

### By Race or Ethnicity

In selected jurisdictions

```{r trend_white}
# White
adv2123_endpoints <- adv2123 %>% 
  filter(group == "White", year == 2022)

p1 <- adv2123 %>% 
  filter(group == "White") %>% 
  ggplot(aes(y = advanced, x = year, color = as.factor(division), text = paste("Division:", division))) +
  geom_line() +
  geom_point() +
  geom_text(data = adv2123_endpoints, size = 4,
            aes(label = division, x = year + 0.3)) + 
  labs(y = "Percent Advanced Diplomas", x = "", title = "Among White Students") +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  scale_x_continuous(breaks = c(2020, 2021, 2022),
                     expand = expansion(add = c(0.05,0.33))) +
  scale_color_manual(values = c("red", "orange", rep("grey",9)),
                     guide = "none") +
  theme_minimal() +
  theme(legend.position = "none")

ggplotly(p1, tooltip = c("text", "y", "x"))
```

```{r trend_black}
# Black
adv2123_endpoints <- adv2123 %>% 
  filter(group == "Black", year == 2022)

p2 <- adv2123 %>% 
  filter(group == "Black") %>% 
  ggplot(aes(y = advanced, x = year, color = as.factor(division), text = paste("Division:", division))) +
  geom_line() +
  geom_point() +
  geom_text(data = adv2123_endpoints, size = 4,
            aes(label = division, x = year + 0.3)) + 
  labs(y = "Percent Advanced Diplomas", x = "", title = "Among Black Students") +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  scale_x_continuous(breaks = c(2020, 2021, 2022),
                     expand = expansion(add = c(0.05,0.33))) +
  scale_color_manual(values = c("red", "orange", rep("grey",9)),
                     guide = "none") +
  theme_minimal() +
  theme(legend.position = "none")

ggplotly(p2, tooltip = c("text", "y", "x"))
```

```{r trend_hisp}
# Hispanic
adv2123_endpoints <- adv2123 %>% 
  filter(group == "Hispanic", year == 2022)

p3 <- adv2123 %>% 
  filter(group == "Hispanic") %>% 
  ggplot(aes(y = advanced, x = year, color = as.factor(division), text = paste("Division:", division))) +
  geom_line() +
  geom_point() +
  geom_text(data = adv2123_endpoints, size = 4,
            aes(label = division, x = year + 0.3)) + 
  labs(y = "Percent Advanced Diplomas", x = "", title = "Among Hispanic Students") +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  scale_x_continuous(breaks = c(2020, 2021, 2022),
                     expand = expansion(add = c(0.05,0.33))) +
  scale_color_manual(values = c("red", "orange", rep("grey",9)),
                     guide = "none") +
  theme_minimal() +
  theme(legend.position = "none")

ggplotly(p3, tooltip = c("text", "y", "x"))
```

```{r trend_asian}
# Asian
adv2123_endpoints <- adv2123 %>% 
  filter(group == "Asian", year == 2022)

p4 <- adv2123 %>% 
  filter(group == "Asian") %>% 
  ggplot(aes(y = advanced, x = year, color = as.factor(division), text = paste("Division:", division))) +
  geom_line() +
  geom_point() +
  geom_text(data = adv2123_endpoints, size = 4,
            aes(label = division, x = year + 0.3)) + 
  labs(y = "Percent Advanced Diplomas", x = "", title = "Among Asian Students") +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  scale_x_continuous(breaks = c(2020, 2021, 2022),
                     expand = expansion(add = c(0.05,0.33))) +
  scale_color_manual(values = c("red", "orange", rep("grey",9)),
                     guide = "none") +
  theme_minimal() +
  theme(legend.position = "none")

ggplotly(p4, tooltip = c("text", "y", "x"))
```

[^1]: For example, the School Quality Profile, College & Career Readiness data for [Albemarle County Public Schools](https://schoolquality.virginia.gov/divisions/albemarle-county-public-schools#desktopTabs-4).