---
title: "Local School Demographics"
author: "Nina Schoonover, Asha Muralidharan, Michele Claibourn, Elizabeth Mitchell, & Christopher Hu"
date: "2024-01-19"
output: 
    html_document:
    template: school-composition-template.html
    css: styles.css
---

```{css, echo=FALSE}
.main-container.container-fluid {
max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
  padding: 0 4.167%;
}
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(tidycensus)
library(reactable)
library(htmltools)
library(sf)
library(sp)
library(tigris)
options(tigris_use_cache = TRUE)
library(leaflet)
library(leaflegend)
library(readxl)

schoolmapdata2023 <- read_csv("data/schoolmapdata_23.csv")

# List of desired localities by FIPS
ccode <- read_csv("data/county_codes.csv")
region <- ccode$code # list of desired counties
# - 003 Albemarle County  
# - 015 Augusta County
# - 029 Buckingham County
# - 540 Charlottesville
# - 065 Fluvanna County
# - 079 Greene County 
# - 109 Louisa County
# - 113 Madison County
# - 125 Nelson County
# - 137 Orange County
# - 790 Staunton
# - 820 Waynesboro

counties_geo <- counties(state = 'VA', year = 2022) # from tigris
counties_geo <- counties_geo %>% subset(COUNTYFP %in% region)

counties_geo <- st_transform(counties_geo, 4326)
# below variables for leaflet map boundary settings
bbox <- st_bbox(counties_geo) %>% as.vector()

```

## Map


```{r}
districtpalette <- colorFactor(
  palette= c("#b45248","#50a9a6","#cab2d6","#2f7e9f","#004600","#360021","#002671", "#70B13D", "#E09056", "#0000CA"),
  domain = c("Albemarle County", "Augusta County", "Buckingham County", "Charlottesville City", "Fluvanna County", "Greene County","Louisa County", "Nelson County", "Staunton City", "Waynesboro City"))

##### add madison and orange to the list above

map <- leaflet() %>% 
  leaflet(width = "100%", height = "700px") %>% 
  addProviderTiles("CartoDB.Voyager") %>% 
  fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>%
  addMapPane('countyBoundaries', zIndex = 410) %>%
  addPolygons(data= counties_geo, color = "#969997",
                    fill = FALSE,
                    weight = 2,
                    group = 'countyBoundaries',
                    options = pathOptions(pane = 'countyBoundaries')) %>%
  addCircleMarkers(data = schoolmapdata2024,
                   lng = ~schoolmapdata2024$Longitude, 
                   lat = ~schoolmapdata2024$Latitude,
                   radius = ~schoolmapdata2024$Total_Count*0.009, 
                   color = ~districtpalette(schoolmapdata2024$District)
                    )

map %>% addLegendFactor(pal = districtpalette,
                shape = "rect",
                orientation = "vertical",
                values = schoolmapdata2024$District,
                naLabel = "",
                title = "",
                opacity = 0.8,
                width = 20,
                height = 10, position = "bottomleft")

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(reactable)

load("2024demographictables.Rdata")

students2024 <- students2024 %>% 
  separate(division, c("division", "locality")) %>% 
  select(-c(year, locality)) %>% 
  relocate(c(count_adv, count_disadv), .before = perc_aian)


```

## Demographic Table {.tabset}

The table below provides the details of the demographic composition of all schools in the school districts in the Piedmont Region/Blue Ridge Health District:

* Albemarle County Public Schools
* Augusta County Public Schools
* Buckingham County Public Schools
* Charlottesville City Schools
* Fluvanna County Public Schools
* Greene County Public Schools
* Louisa County Public Schools
* Nelson County Public Schools
* Staunton City Schools 
* Waynesboro City Schools

According to **Virginia Department of Education (VDOE)** documentation:

* Race or Ethnicity captures "the racial category that most closely reflects the student’s recognition of his or her community or with which the student identifies"

* Economic status "identifies students as economically disadvantaged if they meet any one of the following: 1) is eligible for Free/Reduced Meals, or 2) receives TANF, or 3) is eligible for Medicaid, or 4) identified as either Migrant or experiencing Homelessness."

Scroll further right to see the percent of a school's student body by race and the number and percent of students identified as economically disadvantaged. The tables can be sorted by any column by clicking on the up/down arrow next to the column name.

These numbers reflect Fall Membership numbers provided by the Virginia Department of Education's Fall Membership Build-a-Table site (https://p1pe.doe.virginia.gov/apex_captcha/home.do?apexTypeId=304).

### 2023-2024 School Year
```{r}
reactable(students2024,
          columns = list(
            division = colDef(name = "Division",
                                style = list(background = "#E9F0FC")),
            school = colDef(name = "School",
                                style = list(background = "#E9F0FC")),
            students = colDef(name = "Students",
                                style = list(background = "#E9F0FC")),
            count_aian = colDef(name = "Am. Indian AK Native",
                                style = list(background = "#56B4E9")),
            count_asian = colDef(name = "Asian",
                                style = list(background = "#56B4E9")),
            count_black = colDef(name = "Black",
                                style = list(background = "#56B4E9")),
            count_latinx = colDef(name = "Hispanic",
                                style = list(background = "#56B4E9")),
            count_nhpi = colDef(name = "Native HI Pacific Islander",
                                style = list(background = "#56B4E9")),
            count_white = colDef(name = "White",
                                style = list(background = "#56B4E9")),
            count_multiracial = colDef(name = "Multiracial",
                                style = list(background = "#56B4E9")),
            count_adv = colDef(name = "Advantaged",
                                style = list(background = "#D55E00")),
            count_disadv = colDef(name = "Disadvantaged",
                                style = list(background = "#D55E00")),
            count_nonel = colDef(name = "Non English Learner",
                                style = list(background = "#CC79A7")),
            count_el = colDef(name = "English Learner",
                              style = list(background = "#CC79A7")),
            perc_aian = colDef(name = "Am. Indian AK Native",
                               style = list(background = "#56B4E9"),
                               format = colFormat(suffix = "%")),
            perc_asian = colDef(name = "Asian",
                                style = list(background = "#56B4E9"),
                               format = colFormat(suffix = "%")),
            perc_black = colDef(name = "Black",
                                style = list(background = "#56B4E9"),
                               format = colFormat(suffix = "%")),
            perc_latinx = colDef(name = "Hispanic",
                                style = list(background = "#56B4E9"),
                               format = colFormat(suffix = "%")),
            perc_nhpi = colDef(name = "Native HI Pacific Islander",
                                style = list(background = "#56B4E9"),
                               format = colFormat(suffix = "%")),
            perc_white = colDef(name = "White",
                                style = list(background = "#56B4E9"),
                               format = colFormat(suffix = "%")),
            perc_multiracial = colDef(name = "Multiracial",
                                style = list(background = "#56B4E9"),
                                format = colFormat(suffix = "%")),
            perc_adv = colDef(name = "Advantaged",
                                style = list(background = "#D55E00"),
                               format = colFormat(suffix = "%")),
            perc_disadv = colDef(name = "Disadvantaged",
                                style = list(background = "#D55E00"),
                               format = colFormat(suffix = "%")),
            perc_nonel = colDef(name = "Non English Learner",
                                style = list(background = "#CC79A7"),
                                format = colFormat(suffix = "%")),
            perc_el = colDef(name = "English Learner",
                                style = list(background = "#CC79A7"),
                                format = colFormat(suffix = "%"))),
        columnGroups = list(
            colGroup(name = "Total", 
                     columns = c("division", "school", "students")),
            colGroup(name = "Counts by Racial/Ethnic Identity", 
                     columns = c("count_aian", "count_asian", "count_black", "count_latinx",
                                 "count_nhpi", "count_multiracial", "count_white")),
            colGroup(name = "Counts by Economic Status", 
                     columns = c("count_adv", "count_disadv")),
            colGroup(name = "Counts by English Learner Status", 
                     columns = c("count_el", "count_nonel")),
            colGroup(name = "Percents by Racial/Ethnic Identity", 
                     columns = c("perc_aian", "perc_asian", "perc_black", "perc_latinx",
                                 "perc_nhpi", "perc_multiracial", "perc_white")),
            colGroup(name = "Percents by Economic Status", 
                     columns = c("perc_adv", "perc_disadv")),
            colGroup(name = "Percents by English Learner Status", 
                     columns = c("perc_el", "perc_nonel"))
        ),
          showSortable = TRUE, defaultPageSize = 12)
```

